---
title: 'Chapter 3: Model Now!'
subtitle: 'Loss Exceedance Curves'
author: "Simon Thornewill von Essen"
date: "17 2 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("C:\\Users\\sthornewillvonessen\\Documents\\Online_Courses\\Book_How-To-Measure-Anything-in-Cybersec-Risk")
```

Let's try and do some post processing using the data that we simulated from the last Rmd file.

```{r}
est_loss_n <- dget("./dat/estimate-losses_vector.pkl")
```

Now, instead of plotting a density graph of the losses, we want to think about the probability of loss exceedance.

```{r}
min_loss <- 0
max_loss <- max(est_loss_n)

min_loss
max_loss
```

So, now we can calculate our vector of losses

```{r}
x_loss <- seq(min_loss, max_loss, 200)

head(x_loss)
```

Now, we want to calculate the probability

```{r}
n.vals <- length(est_loss_n)
p_loss <- rep(0, length(x_loss))

for(i in 1:length(p_loss)){
  p <- sum(est_loss_n > x_loss[i])/n.vals
  p_loss[i] <- p
}

plot(x_loss, p_loss, type="l", col="red")
```

Now that we have plotted our LEC, we can consider what kind of loss we are willing to tolerate. We can plot a loss excedance tolerance that shows what level of risk we are willing to take on.

The LET is something that we can calculate for ourselves, by considering multiple losses and assigning a desired probability of occurance to them, we can then plot this line and compare it with the inherent loss in our LEC.
